Methods and Results

### 2022-05-03 ms.Rmd
Updated paper woth diagnostic checks and some more nuanced results base don those checks as there appears to be a big outlier in one model at least.

  model1.1 <- rma.mv(SMDH ~ 1, V = data$v_SMDH, 
                     random = list(~1|study, ~1|variable, ~1|species, ~1|obs), test = "t", data = data)
    robust <- robust(model1.1, cluster = data$study, vcov = "CR2")

  model1.2 <- rma.mv(SMDH ~ 1, V = data$v_SMDH, 
                     random = list(~1|study, ~1|variable, ~1|species, ~1|species2, ~1|obs), R = list(species2 = phylogeny), test = "t", data = data)

@@

@@

       model1.1_pub <- rma.mv(SMDH ~ sqrt(inv_n_eff), V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~1|obs), test = "t", data = data)

  model1.1_pub_full <- rma.mv(SMDH ~ sqrt(inv_n_eff) + SSD_lnRR + Harm_type, V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~1|obs), test = "t", data = data)
   model1.1_pub_full_het <- rma.mv(SMDH ~ sqrt(inv_n_eff) + SSD_lnRR + Harm_type, V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~Harm_type|obs), rho = 0, 
                              str = "HCS", test = "t", data = data)
      AICc(model1.1_pub_full, model1.1_pub_full_het) # Basically the same results so simplify model.

###################
# Meta-regressions
###################
  ###############
      # SSD
  ################
  # Marginalised model. Just easier to get slope
  model1.3 <- rma.mv(SMDH ~ zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data)
  robust_model1.3 <- robust(model1.3, cluster = data$study)

  # Use raw SSD for plotting, so we can plot raw data, which is easier to interpret
  model1.3_nonZ <- rma.mv(SMDH ~ SSD_lnRR, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data)
  robust_model1.3_nonZ <- robust(model1.3_nonZ, cluster = data$study)

@@

  ###############
      # Harm type
  ################
  # Marginalised Harm type        
  model1.4 <- rma.mv(SMDH ~ Harm_type, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data)
  robust(model1.4, cluster = data$study)

@@

  ###############
      # Harm type & SSD
  ################
  # Model that contains harm type and SSD. Harm type appears to have different variances, so I'll check whether it is safe to assume homogeneity of variance across groups
  model1.5 <- rma.mv(SMDH ~  Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                     method = "ML", data = data)
  model1.5_inter <- rma.mv(SMDH ~  Harm_type + zSSD + Harm_type:zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                     method = "ML", data = data)

@@

  # Non Z transform
  model1.5_nonZ <- rma.mv(SMDH ~ Harm_type + SSD_lnRR, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data)
  robust_model1.5_nonZ <- robust(model1.5_nonZ, cluster = data$study)

  model1.6 <- rma.mv(SMDH ~ Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~(1 + Harm_type)|obs, ~1|species), 
                     rho = 0, struc = "HCS", test = "t", data = data)

@@

  # Check robustness of results using sandwich estimator
  robust_model1.5 <- robust(model1.5, cluster = data$study, vcov = "CR2")

@@

  model1.7_inetr <- rma.mv(SMDH ~  SCR.SCI*zSSD, V = sperm_data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                     method = "ML", data = sperm_data)

  model1.7 <- rma.mv(SMDH ~  SCR.SCI + zSSD, V = sperm_data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                     method = "ML", data = sperm_data)

@@

@@

  ###############
      # Lifespan
  ################
  # Sex differences in lifespan
    model1.8 <- rma.mv(SMDH ~ sex_life, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                       method = "REML", data = data)
    robust_model1.8 <- robust(model1.8, cluster = data$study, vcov = "CR2")

    model1.9 <- rma.mv(SMDH ~ sex_life + zSSD + Harm_type, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                       method = "REML", data = data)
    robust_model1.9 <- robust(model1.8, cluster = data$study, vcov = "CR2")

  # Female lifespan
    model1.10 <- rma.mv(SMDH ~ scale(F_LRS), V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                       method = "REML", data = data)
    robust_model1.10 <- robust(model1.9, cluster = data$study, vcov = "CR2")

    model1.11 <- rma.mv(SMDH ~ scale(F_LRS) + Harm_type, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), test = "t", 
                       method = "REML", data = data)
    robust_model1.11 <- robust(model1.11, cluster = data$study, vcov = "CR2")

@@

Experimentally manipulating female harm resulted in a strong decrease in female fitness overall (i.e., positive effect size with control group females having higher fitness than treatment groups: `r model1.1$beta`, 95% CI: `r model1.1$ci.lb` to `r model1.1$ci.ub`, $n_{effects}$ = `r model1.1$k.all`, $n_{study}$ = `r model1.1$s.nlevels[1]`). This strong effect held even when accounting for within-study non-independence (Meta-analytic mean using Robust Variance Estimator: `r robust$beta`, 95% CI: `r robust$ci.lb` to `r robust$ci.ub`). Visual inspection of funnel plot asymmetry suggested evidence for publication bias (i.e., missed effects sizes with low precision when female harm was worse in control treatments) (Figure \@ref(fig:funnel)), which was confirmed through the identification of a significant slope between effect size and effective sample size ($\beta$ = `r model1.1_pub$beta[2]`, 95% CI: `r model1.1_pub$ci.lb[2]` to `r model1.1_pub$ci.ub[2]`). This result held true when accounting for heterogeneity using meta-regression models ($\beta$ =`r model1.1_pub_full$beta[2]`, 95% CI: `r model1.1_pub_full$ci.lb[2]` to `r model1.1_pub_full$ci.ub[2]`). Correcting for the possibility of missing studies resulted in the overall meta-analytic mean effect size being indistinguishable from zero (Corrected meta-analytic mean: `r model1.1_pub$beta[1]`, 95% CI: `r model1.1_pub$ci.lb[1]` to `r model1.1_pub$ci.ub[1]`). These findings suggest that additional empirical studies may change the overall magnitude of effects, possibly showing weaker effects then what is currently published, however, it is important to recognise that we cannot know for sure what (if anything) is generating such a bias.

@@

#### Negative effects on fitness depend on the type of harm done to females

Male harm type impacted the overall magnitude of effects explaining `r male_harm_r2[1]*100`% of effect size variance (Figure \@ref(fig:maleharm)). Species where mating results in both direct and indirect harm to females have males that significantly impact female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[1]`, 95% CI: `r robust_model1.5_2$ci.lb[1]` to `r robust_model1.5_2$ci.ub[1]`,  p = `r p_value(robust_model1.5_2$pval[1])`). Surprisingly, however, species with direct female harm only (i.e., traumatic insemination), showed a small and opposite effect on female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[2]`, 95% CI: `r robust_model1.5_2$ci.lb[2]` to `r robust_model1.5_2$ci.ub[2]`; Figure \@ref(fig:maleharm)), however, this effect did not differ significantly from zero (p = `r p_value(robust_model1.5_2$pval[2])`). While indirect male harm also resulted in a smaller overall meta-analytic mean relative to species with both, it was still significantly positive (overall meta-analytic mean: `r robust_model1.5_2$beta[3]`, 95% CI: `r robust_model1.5_2$ci.lb[3]` to `r robust_model1.5_2$ci.ub[3]`,  p = `r p_value(robust_model1.5_2$pval[3])`), and only marginally different than species with both direct and indirect male harm (Figure \@ref(fig:maleharm)).

```{r maleharm, fig.width=11.56944, fig.height=7.62500, fig.cap="Distribution of effect sizes, overall meta-analytic mean estimates and 95% confidence intervals (CIs) across species with different types of male harm towards females. Total effect sizes (n) and total number of studies (k) are provided for each level of Male Harm Type. Relevant contrasts between meta-analytic means are provided, along with 95% CIs and significance of contrast."}

@@

>>>>

  model1.1 <- rma.mv(SMDH ~ 1, V = data$v_SMDH, 
                     random = list(~1|study, ~1|variable, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data)
    robust <- robust(model1.1, cluster = data$study, vcov = "CR2")

  model1.2 <- rma.mv(SMDH ~ 1, V = data$v_SMDH, 
                     random = list(~1|study, ~1|variable, ~1|species, ~1|species2, ~1|obs), R = list(species2 = phylogeny),  dfs = "contain", test = "t", data = data)

@@

        # Just some model checks
         checks = FALSE

         if(checks){
           hist(residuals(model1.1))

           cd.model1.1 <- cooks.distance(model1.1)
           plot.cook(cd.model1.1, model1.1, vals = "obs")

           vals <- plot.cook(cd.model1.1, model1.1, vals = "obs", type = "vals")

           # results highly robust to outliers removed. Effect size strength increases
           model1.1_refit <- rma.mv(SMDH ~ 1, V = data$v_SMDH, 
                     random = list(~1|study, ~1|variable, ~1|species, ~1|obs), test = "t", data = data[!data$obs %in% vals,])
         }

@@

       model1.1_pub <- rma.mv(SMDH ~ sqrt(inv_n_eff), V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data)

  model1.1_pub_full <- rma.mv(SMDH ~ sqrt(inv_n_eff) + SSD_lnRR + Harm_type, V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~1|obs), dfs = "contain", test = "t", data = data)
   model1.1_pub_full_het <- rma.mv(SMDH ~ sqrt(inv_n_eff) + SSD_lnRR + Harm_type, V = data$v_SMDH, 
                              random = list(~1|study, ~1|variable, ~1|species, ~Harm_type|obs), rho = 0, 
                              str = "HCS",  dfs = "contain", test = "t", data = data)
      AICc(model1.1_pub_full, model1.1_pub_full_het) # Basically the same results so simplify model.

###################
# Meta-regressions
###################
  ###############
      # SSD
  ################
  # Marginalised model. Just easier to get slope
  model1.3 <- rma.mv(SMDH ~ zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data[complete.cases(data$zSSD),])
  robust_model1.3 <- robust(model1.3, cluster = data[complete.cases(data$zSSD),"study"])

  # Use raw SSD for plotting, so we can plot raw data, which is easier to interpret
  model1.3_nonZ <- rma.mv(SMDH ~ SSD_lnRR, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data)
  robust_model1.3_nonZ <- robust(model1.3_nonZ, cluster = data$study)

@@

  # Just some model checks.
    if(checks){
           hist(residuals(model1.3))

           cd.model1.3 <- cooks.distance(model1.3)
           plot.cook(cd.model1.3, model1.3, vals = "obs")
           vals <- plot.cook(cd.model1.3, model1.3, vals = "obs", type = "vals")

           # Removing outliers decrease zSSD effect by 0.10, but still non-significant/
           model1.3_refit <- rma.mv(SMDH ~ zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data[!data$obs %in% vals,])
    }

  ###############
      # Harm type
  ################
  # Marginalised Harm type        
  model1.4 <- rma.mv(SMDH ~ Harm_type, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data[complete.cases(data$Harm_type),])
  robust(model1.4, cluster = data[complete.cases(data$Harm_type),"study"])

@@

  # Just some model checks.
    if(checks){
           hist(residuals(model1.4))

           cd.model1.4 <- cooks.distance(model1.4)
           plot.cook(cd.model1.4, model1.4, "obs")
           vals <- plot.cook(cd.model1.4, model1.4, "obs", type = "vals")

         model1.4_refit <- rma.mv(SMDH ~ Harm_type, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs),  dfs = "contain", test = "t", data = data[!data$obs %in% vals,])

        }
  ###############
      # Harm type & SSD
  ################
  # Model that contains harm type and SSD. Harm type appears to have different variances, so I'll check whether it is safe to assume homogeneity of variance across groups
  model1.5 <- rma.mv(SMDH ~  Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species),  dfs = "contain", test = "t", 
                     method = "ML", data = data[complete.cases(data$Harm_type, data$zSSD),])
  model1.5_inter <- rma.mv(SMDH ~  Harm_type + zSSD + Harm_type:zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species),  dfs = "contain", test = "t", 
                     method = "ML", data = data)

@@

  # Non Z transform
  model1.5_nonZ <- rma.mv(SMDH ~ Harm_type + SSD_lnRR, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), dfs = "contain", test = "t", data = data)
  robust_model1.5_nonZ <- robust(model1.5_nonZ, cluster = data$study)

  model1.6 <- rma.mv(SMDH ~ Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~(1 + Harm_type)|obs, ~1|species), 
                     rho = 0, struc = "HCS",  dfs = "contain", test = "t", data = data)

@@

  # Check robustness of results using sandwich estimator
  robust_model1.5 <- robust(model1.5, cluster = data[complete.cases(data$Harm_type, data$zSSD),"study"], vcov = "CR2")

  # Just some model checks.
   if(checks){
           hist(residuals(model1.5))

           cd.model1.5 <- cooks.distance(robust_model1.5)
           plot.cook(cd.model1.5, robust_model1.5, "obs")
           vals <- plot.cook(cd.model1.5, robust_model1.5, "obs", type = "vals")

         model1.5_refit <- rma.mv(SMDH ~ Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data[!data$obs %in% vals,])    
         # But very clear outlier is 136 so remove this first
         model1.5_refit <- rma.mv(SMDH ~ Harm_type + zSSD, V = data$v_SMDH, 
                     random = list(~1|study, ~1|species, ~1|obs), test = "t", data = data[!data$obs %in% 136,])    
   }

@@

  model1.7_inetr <- rma.mv(SMDH ~  SCR.SCI*zSSD, V = sperm_data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                     method = "ML", data = sperm_data)

  model1.7 <- rma.mv(SMDH ~  SCR.SCI + zSSD, V = sperm_data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                     method = "ML", data = sperm_data)

@@

  # Just some model checks.
    if(checks){
           hist(residuals(model1.7))

           cd.model1.7 <- cooks.distance(model1.7)
           plot.cook(cd.model1.7, model1.7, "obs")
           obs <- plot.cook(cd.model1.7, model1.7, "obs",type = "vals")

      # Results are pretty similar without these rows with high Cooks D
      model1.7_refit <- rma.mv(SMDH ~  SCR.SCI + zSSD, V = sperm_data$v_SMDH, 
                     random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                     method = "ML", data = sperm_data[!sperm_data$obs %in% obs,])
        }

@@

  ###############
      # Lifespan
  ################
  # Sex differences in lifespan
    model1.8 <- rma.mv(SMDH ~ sex_life, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                       method = "REML", data = data)
    robust_model1.8 <- robust(model1.8, cluster = data$study, vcov = "CR2")

    model1.9 <- rma.mv(SMDH ~ sex_life + zSSD + Harm_type, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                       method = "REML", data = data)
    robust_model1.9 <- robust(model1.9, cluster = data$study, vcov = "CR2")

  # Female lifespan
    model1.10 <- rma.mv(SMDH ~ scale(F_LRS), V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                       method = "REML", data = data)
    robust_model1.10 <- robust(model1.10, cluster = data$study, vcov = "CR2")

    model1.11 <- rma.mv(SMDH ~ scale(F_LRS) + Harm_type + zSSD, V = data$v_SMDH, 
                       random = list(~1|study, ~1|obs, ~1|species), dfs = "contain", test = "t", 
                       method = "REML", data = data)
    robust_model1.11 <- robust(model1.11, cluster = data$study, vcov = "CR2")

@@

Experimentally manipulating female harm resulted in a strong decrease in female fitness overall (i.e., positive effect size with control group females having higher fitness than treatment groups: `r model1.1$beta`, 95% CI: `r model1.1$ci.lb` to `r model1.1$ci.ub`, $n_{effects}$ = `r model1.1$k.all`, $n_{study}$ = `r model1.1$s.nlevels[1]`). This effect held even when accounting for within-study non-independence (meta-analytic mean using Robust Variance Estimator: `r robust$beta`, 95% CI: `r robust$ci.lb` to `r robust$ci.ub`). Visual inspection of funnel plot asymmetry suggested evidence for publication bias (i.e., missed effects sizes with low precision when female harm was worse in control treatments) (Figure \@ref(fig:funnel)), which was confirmed through the identification of a significant slope between effect size and effective sample size ($\beta$ = `r model1.1_pub$beta[2]`, 95% CI: `r model1.1_pub$ci.lb[2]` to `r model1.1_pub$ci.ub[2]`). This result held true when accounting for heterogeneity using meta-regression models ($\beta$ =`r model1.1_pub_full$beta[2]`, 95% CI: `r model1.1_pub_full$ci.lb[2]` to `r model1.1_pub_full$ci.ub[2]`). Correcting for the possibility of missing studies resulted in the overall meta-analytic mean effect size being indistinguishable from zero (Corrected meta-analytic mean: `r model1.1_pub$beta[1]`, 95% CI: `r model1.1_pub$ci.lb[1]` to `r model1.1_pub$ci.ub[1]`). These findings suggest that additional empirical studies may change the overall magnitude of effects, possibly showing weaker effects then what is currently published, but our current state of knowledge suggests that female fitness is indeed compromised overall by male harm.

@@

#### Negative effects on fitness and the type of harm done to females

Male harm type appeared to impact the overall magnitude of effects explaining `r male_harm_r2[1]*100`% of effect size variance (Figure \@ref(fig:maleharm)). Species where mating results in both direct and indirect harm to females have males that significantly impact female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[1]`, 95% CI: `r robust_model1.5_2$ci.lb[1]` to `r robust_model1.5_2$ci.ub[1]`,  p = `r p_value(robust_model1.5_2$pval[1])`). Surprisingly, however, species with direct female harm only (i.e., traumatic insemination), showed a small and opposite effect on female fitness (overall meta-analytic mean: `r robust_model1.5_2$beta[2]`, 95% CI: `r robust_model1.5_2$ci.lb[2]` to `r robust_model1.5_2$ci.ub[2]`; Figure \@ref(fig:maleharm)), however, this effect did not differ significantly from zero (p = `r p_value(robust_model1.5_2$pval[2])`). While indirect male harm also resulted in a smaller overall meta-analytic mean relative to species with both, it was still significantly positive (overall meta-analytic mean: `r robust_model1.5_2$beta[3]`, 95% CI: `r robust_model1.5_2$ci.lb[3]` to `r robust_model1.5_2$ci.ub[3]`,  p = `r p_value(robust_model1.5_2$pval[3])`), and only marginally different than species with both direct and indirect male harm (Figure \@ref(fig:maleharm)). Regardless, differences among species exhibiting different types of female harm seemed to be driven primarily by
a single study (Taylor, 2008) on *Drosophila melanogaster* (very high Cook's distance). Removing this effect resulted in no significant difference between harm type categories.

```{r maleharm, fig.width=11.56944, fig.height=7.62500, fig.cap="Distribution of effect sizes, overall meta-analytic mean estimates and 95% confidence intervals (CIs) across species with different types of male harm towards females. Total effect sizes (n) and total number of studies (k) are provided for each level of Male Harm Type. Relevant contrasts between meta-analytic means are provided, along with 95% CIs and significance of contrast. Note that this includes all data. Removing outlier point from single study results in no different between harm type categories. "}

@@

#### Species lifespan does not explain patterns in female harm
Female lifespan or the difference in lifespan between males and females of a species did not explain significant variation in female harm in treatments after controlling for sexual size dimorphism and harm type ($\beta_{lifespan_{f}}$ = `r robust_model1.9$b[2]`, 95% CI: `r robust_model1.9$ci.lb[2]` to `r robust_model1.9$ci.ub[2]`, t = `r robust_model1.9$zval[2]`, df = `r robust_model1.9$dfs`, p = `r robust_model1.9$pval[2]`; $\beta_{lifespan_{\Delta_{m-f}}}$ = `r robust_model1.11$b[2]`, 95% CI: `r robust_model1.11$ci.lb[2]` to `r robust_model1.11$ci.ub[2]`, t = `r robust_model1.11$zval[2]`, df = `r robust_model1.11$dfs`, p = `r robust_model1.11$pval[2]`)
